---
title: "ndvi"
format: html
editor: visual
---

# NDVI Stuff

Step Overview:

1.  Download raster data

    a.  NAIP "year_c" files
    b.  elevation data
    c.  SCWRP vegetation data

2.  Conduct NDVI analysis ((band_4 - band_3)/(band_4 + band_3)) (it might be flipped)

3.  Add NDVI values to vegetation dataframe

4.  Run a correlation analysis

    a.  %cover \~ NDVI, veg_height \~ NDVI

# HeadR (get it)

Load packages 😏

```{r}
library(tidyverse)
library(dplyr)
library(readxl)
library(scales)
library(sjPlot)
library(lmerTest)
library(grateful)
library(rtide)
library(weathermetrics)
```

Import excel sheet

```{r}
cord_data <- read_excel("cordgrass_tings.xlsx")
```

Get data ready for analysis (wash your hands, get a drink, grab a plate etc.)

```{r}
cord_data <- cord_data |>
  mutate(pct_cover = ((VegetatedCover/100) * (EstimatedCover))) |>
  mutate(Station = as.character(Station))

cord_data
```

### graphs for brownie points

```{r}
ggplot(cord_data) +
  aes(NDVI, VegetatedCover, colour = Tide) +
  geom_point() +
  scale_color_manual(values = c("green", "red2", "darkorange")) +
  theme_grey(base_size = 14)
```

```{r}
ggplot(cord_data) +
  aes(NDVI, VegetatedCover, colour = Habitat) +
  geom_point() +
  scale_color_manual(values = c("green", "red2", "darkorange")) +
  theme_grey(base_size = 14)

ggplot(cord_data) +
  aes(NDVI, TallestPlantHeight, colour = Station) +
  geom_point() +
  scale_color_manual(values = c("green", "red2", "darkorange")) +
  theme_grey(base_size = 14)
```

## My data is sooooo normal (spoiler alert: NOT)

```{r}
cord_model1 <- lm(pct_cover ~ NDVI, data = cord_data)
cord_model2 <- aov(pct_cover ~ NDVI, data = cord_data)

residuals_1 <- residuals(cord_model2)

#normie test
shapiro.test(residuals_1)


plot(cord_model1, which = 1)
plot(cord_model1, which = 2)
```

### why cant you be normal

```{r}
cord_model3 <- lm(TallestPlantHeight ~ NDVI, data = cord_data)
cord_model4 <- aov(TallestPlantHeight ~ NDVI, data = cord_data)

residuals_2 <- residuals(cord_model4)

#normie test
shapiro.test(residuals_2)

plot(cord_model3, which = 1)
plot(cord_model3, which = 2)

anova(cord_model1)
anova(cord_model3)
```

\########################################################################

# This is where the fun begins

\########################################################################

## Start of four estuary analysis

### 1. Load data!

```{r}
combined_data <- read_excel("four_ndvi_inc.xlsx")


```

### 2. Prep data!

```{r}
combined_data <- combined_data |>
  mutate(stationno = as.character(stationno)) |>
  mutate(raw_total_stems = ifelse(raw_total_stems < 0, 0, raw_total_stems))

```

### 3. Preliminary graphs!

```{r}
## first three by station, second three by tide!
ggplot(combined_data) +
  aes(NDVI, veg_cov_pct, colour = stationno) +
  geom_point() +
  scale_color_manual(values = c("green", "red2", "darkorange")) +
  theme_grey(base_size = 14)

ggplot(combined_data) +
  aes(NDVI, meanCGht, colour = stationno) +
  geom_point() +
  scale_color_manual(values = c("green", "red2", "darkorange")) +
  theme_grey(base_size = 14)

ggplot(combined_data) +
  aes(NDVI, maxCGht, colour = stationno) +
  geom_point() +
  scale_color_manual(values = c("green", "red2", "darkorange")) +
  theme_grey(base_size = 14)

#####

ggplot(combined_data) +
  aes(NDVI, veg_cov_pct, colour = tide) +
  geom_point() +
  scale_color_manual(values = c("green", "red2", "darkorange")) +
  theme_grey(base_size = 14)

ggplot(combined_data) +
  aes(NDVI, meanCGht, colour = tide) +
  geom_point() +
  scale_color_manual(values = c("green", "red2", "darkorange")) +
  theme_grey(base_size = 14)

ggplot(combined_data) +
  aes(NDVI, maxCGht, colour = tide) +
  geom_point() +
  scale_color_manual(values = c("green", "red2", "darkorange")) +
  theme_grey(base_size = 14)


ggplot(combined_data) +
  aes(x = stem_density) +
  geom_histogram()


```

Mean and max cordgrass height (figures 5 and 6) look like there might be something interesting there. here's hoping!

### 4. Check for normality!

Plots for vegetation cover percentage

```{r}
combo_model1 <- lm(veg_cov_pct ~ NDVI, data = combined_data)
combo_model2 <- aov(veg_cov_pct ~ NDVI, data = combined_data)

residuals_1 <- residuals(combo_model2)

#normie test
shapiro.test(residuals_1)


plot(combo_model1, which = 1)
plot(combo_model1, which = 2)

anova(combo_model1)
```

Coverage percentage is significant! What about cordgrass coverage?

```{r}
combo_model3 <- lm(estimatedcover ~ NDVI, data = combined_data)
combo_model4 <- aov(estimatedcover ~ NDVI, data = combined_data)

residuals_2 <- residuals(combo_model4)

#normie test
shapiro.test(residuals_2)


plot(combo_model3, which = 1)
plot(combo_model3, which = 2)

anova(combo_model3)
```

Plots for max cordgrass height

```{r}
combo_model5 <- lm(maxCGht ~ NDVI, data = combined_data)
combo_model6 <- aov(maxCGht ~ NDVI, data = combined_data)

residuals_3 <- residuals(combo_model6)

#normie test
shapiro.test(residuals_3)

plot(combo_model5, which = 1)
plot(combo_model5, which = 2)

anova(combo_model5)
```

Plots for mean cordgrass height

```{r}
combo_model7 <- lm(meanCGht ~ NDVI, data = combined_data)
combo_model8 <- aov(meanCGht ~ NDVI, data = combined_data)

residuals_4 <- residuals(combo_model8)

#normie test
shapiro.test(residuals_4)


plot(combo_model7, which = 1)
plot(combo_model7, which = 2)

anova(combo_model7)
```

Plots for total stems

```{r}
combo_model9 <- lm(raw_total_stems ~ NDVI, data = combined_data)
combo_model10 <- aov(raw_total_stems ~ NDVI, data = combined_data)

residuals_5 <- residuals(combo_model10)

#normie test
shapiro.test(residuals_5)


plot(combo_model9, which = 1)
plot(combo_model9, which = 2)

anova(combo_model7)
```

### 5. Let's try some bigger linear models!

```{r}
library(lme4)
library(lmerTest)
library(MuMIn)
library(cAIC4)
```

```{r}
lmer_model1 <- lmer(NDVI ~ maxCGht + raw_total_stems + (1|estuary), data = combined_data)
summary(lmer_model1)
```

```{r}
lmer_model3 <- lmer(NDVI ~ maxCGht * raw_total_stems + (1|estuary), data = combined_data)
summary(lmer_model3)
```

```{r}
lmer_model2 <- lmer(NDVI ~ maxCGht + raw_total_stems + (1|tide) + (1|estuary), data = combined_data)
summary(lmer_model2)
```

```{r}
lmer_model4 <- lmer(NDVI ~ maxCGht * raw_total_stems + (1|tide) + (1|estuary), data = combined_data)
summary(lmer_model4)
```

```{r}
lmer_model5 <- lmer(NDVI ~ maxCGht + raw_total_stems + stem_density + (1|tide) + (1|estuary), data = combined_data)
summary(lmer_model5)
```

```{r}
lmer_model6 <- lmer(NDVI ~ maxCGht * raw_total_stems * stem_density + (1|tide) + (1|estuary), data = combined_data)
summary(lmer_model6)
```

```{r}
lmer_model7 <- lmer(NDVI ~ maxCGht * raw_total_stems + stem_density + (1|tide) + (1|estuary), data = combined_data)
summary(lmer_model7)
```

```{r}
lmer_model8 <- lmer(NDVI ~ maxCGht + raw_total_stems + estimatedcover + (1|tide) + (1|estuary), data = combined_data)
summary(lmer_model8)
```

```{r}
lmer_model9 <- lmer(NDVI ~ maxCGht * raw_total_stems * stem_density * estimatedcover + (1|tide) + (1|estuary), data = combined_data)
summary(lmer_model9)
```

```{r}
lmer_model10 <- lmer(NDVI ~ maxCGht * raw_total_stems * estimatedcover + (1|tide) + (1|estuary), data = combined_data)
summary(lmer_model10)
```

```{r}
lmer_model11 <- lmer(NDVI ~ maxCGht + raw_total_stems + stem_density + estimatedcover + (1|tide) + (1|estuary), data = combined_data)
summary(lmer_model11)
```

```{r}
lmer_model12 <- lmer(NDVI ~ estimatedcover + raw_total_stems + (1|estuary), data = combined_data)
summary(lmer_model12)
```

```{r}
lmer_model13 <- lmer(NDVI ~ estimatedcover * raw_total_stems + (1|estuary), data = combined_data)
summary(lmer_model13)
```

```{r}
lmer_model14 <- lmer(NDVI ~ estimatedcover + raw_total_stems + (1|tide) + (1|estuary), data = combined_data)
summary(lmer_model14)
```

#### 5.a) cAIC comparison!

```{r}
#AIC comparison
AIC1 <- cAIC(lmer_model1)
AIC2 <- cAIC(lmer_model2)
AIC3 <- cAIC(lmer_model3)
AIC4 <- cAIC(lmer_model4)
AIC5 <- cAIC(lmer_model5)
AIC6 <- cAIC(lmer_model6)
AIC7 <- cAIC(lmer_model7)
AIC8 <- cAIC(lmer_model8)
AIC9 <- cAIC(lmer_model9)
AIC10 <- cAIC(lmer_model10)
AIC11 <- cAIC(lmer_model11)
AIC12 <- cAIC(lmer_model12)
AIC13 <- cAIC(lmer_model13)
AIC14 <- cAIC(lmer_model14)

AIC1
AIC2
AIC3
AIC4
AIC5
AIC6
AIC7
AIC8
AIC9
AIC10
AIC11
AIC12
AIC13
AIC14
```

```{r}
models_list <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14")
caic <- c(172.78, 180.04, 172.55,  180.04, 181.15, 183.41, 180.93, 178.02, 169.12, 177.39, 179.68, 168.20, 166.30, 176.66)
caic_data <- data.frame(models = models_list, cAIC = caic)
caic_data$models <- factor(caic_data$models, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14"))

lmer_plot <- ggplot(caic_data, aes(x = models, y = cAIC)) +
  geom_bar(stat = "identity", fill = "cornflowerblue", width = 0.5) +
  geom_text(aes(label = cAIC), vjust = -0.5, size = 3) +
  labs(x = "Model Iteration", y = "cAIC Value")

lmer_plot


##showing these three due to their lower cAIC values
ggpubr::ggqqplot(residuals(lmer_model9))
ggpubr::ggqqplot(residuals(lmer_model12))
ggpubr::ggqqplot(residuals(lmer_model13))

#it seems that model 12 might be a slightly better fit?



plot_model(lmer_model9) + geom_hline(aes(yintercept = 0), linetype = "dotted")

plot_model(lmer_model12) + geom_hline(aes(yintercept = 0), linetype = "dotted")

plot_model(lmer_model13) + geom_hline(aes(yintercept = 0), linetype = "dotted")
```

### 6. GLM time!

```{r}
glm_1 <- glm(raw_total_stems ~ NDVI, data = combined_data, family = "poisson")
summary(glm_1)
```

```{r}
predicted_count <- predict(glm_1, type = "response")
glm_plot <- data.frame(observed = combined_data$raw_total_stems, predicted = predicted_count, NDVI = combined_data$NDVI)

glm_plot |>
  ggplot(aes(x = NDVI, y = observed)) +
  geom_point(color = "cornflowerblue", alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "poisson"), aes(y = predicted), color = "red4", se = FALSE) +
  labs(y = "total stems")

glm_plot
  
```

In my completely professional opinion, it seems that the total \# of stems decreases slightly as NDVI increases because of the transect proximity to water. Values around 0.05 to 0.2 usually indicate bare ground, while -.1 to 0 usually correspond to sandy areas.
