---
title: "TidesDemo for Advanced Quant"
author: "Amy Henry"
date: "3/2/2023"
output: html_document
---

Load packages, including rtide, which does tide predictions
https://cran.r-project.org/web/packages/rtide/index.html

```{r}
#install.packages("weathermetrics")
#install.packages("rtide")

library(rtide)
library(dplyr)
library(lubridate)
library(tidyverse)
```

This is data I collected with some HOBO dataloggers during my thesis at a beach. The loggers were placed intertidally, so sometimes they were in the air and sometimes in the water. There were two loggers placed recording temperatures, HI (higher in the intertidal) and LO (lower in the intertidal)
```{r}
tempdata <- read.csv(file = "FieldHoboTemp2013_jun3.csv")
tempdata
```

Back in 2017, before lubridate was a thing, we had to use these terrible date import formats. This chunk just gets my date data into the proper time zone, including a conversion for daylight savings time. 
```{r}
DateTime <- strptime(tempdata$Date.Time..GMT.07.00, "%m/%d/%y %H:%M", tz = "PST8PDT") #tz is for timezone Pacific
DateTime <- as.POSIXct(DateTime) #formatting crap
head(DateTime)
```

```{r}
glimpse(tempdata)

DateTime_tidy <- tempdata %>% mutate(DateTime =  mdy_hm(Date.Time..GMT.07.00))



```




The function tide_stations is from the rtide package and pulls data from the location of your choice. 
```{r}
# Friday <- tide_stations(".*Friday", harmonics = rtide::harmonics) #selects tide station for Friday Harbor
# Station <- rep(Friday, nrow(tempdata)) #repeats it once for each row of the temperature dat
# harmonicobject <- data_frame(Station, DateTime) # makes a new frame with the DateTimes from my temperature data plus the station name repeated for each row

# FHtides <- tide_height_data(harmonicobject, harmonics = rtide::harmonics) #compute the tide at the time
```

```{r}
Newport <- tide_stations(".*Corona", harmonics = rtide::harmonics) #selects tide station for Newport Bay Entrance at Corona del Mar

Station <- rep(Newport, nrow(tempdata)) #repeats it once for each row of the temperature data

harmonicobject <- tibble(Station, DateTime_tidy) # makes a new frame with the DateTimes from my temperature data plus the station name repeated for each row

NewportTides <- tide_height_data(harmonicobject, harmonics = rtide::harmonics) #compute the tide at the time
```

These lines bind the temp data to the tide data and then do some formatting junk
```{r}
tempdata <- cbind(tempdata, FHtides)

samedates <- tempdata %>% drop_na(HI, LO)

# Are HI and Lo different 
t.test(samedates$HI,samedates$LO, paired = TRUE)
plot(density(tempdata$HI, na.rm = TRUE))

samedates <- samedates %>% mutate(HI = weathermetrics::fahrenheit.to.celsius(HI),
                                     LO = weathermetrics::fahrenheit.to.celsius(LO))

```

This sets up the plot, which is made in R basic instead of ggplot, but I am very proud of it. 
```{r}
par(mfrow = c(1,1), fig = c(0, 0.9, 0, 0.77))
#par(, new = TRUE)
plot(samedates$HI~samedates$DateTime, type = "l", ylab = "Temperature, degrees C", xlab = "Date",
     #main = "Temperatures at Crescent Beach, Orcas Island, WA, June - August 2013"
     #ylim = c(50, 85), 
     bty = "l", cex.lab = 0.7, cex.axis = 0.6,
     xaxt="n")
axis.POSIXct(1, at=seq(min(samedates$DateTime), max(samedates$DateTime), by = "3 day"), 
           format="%m-%d", cex.axis = 0.7)

lines(samedates$LO~samedates$DateTime, type = "l", col = "red")
#axis(side = 1, at = c("2013-06-10 00:00:00","2013-06-20 00:00:00","2013-06-30 00:00:00",
#                      "2013-07-10 00:00:00","2013-07-20 00:00:00","2013-07-30 00:00:00",
#                      "2013-08-10 00:00:00","2013-08-20 00:00:00","2013-08-30 00:00:00"),
#     labels = c("June 10", "June 20", "June 30", "July 10", "July 20", "July 30", "Aug 10", 
#                "Aug 20","Aug 30"))
#abline(h = 90)
#par(new = T)
#rect(0,1, 0, 1,density = NULL, col = "white", border = "green")
par(fig = c(0, 0.9, 0.45, 1), new = TRUE)
plot(samedates$TideHeight~samedates$DateTime, type = "l",bty = "l", 
     axes=F, 
     xlab=NA, ylab=NA, col = "blue", cex.lab = 0.7, cex.axis = 0.6)
abline(h = -0.015, lty = "dashed", col = "black")
abline(h = -0.304, lty = "dashed", col = "red")
axis(side = 2, cex.axis = 0.6)
mtext(side = 2, line = "3", 'Tide Height, m', cex = 0.7)
#legend("topleft",
#       legend=c(expression(-log[10](italic(p))), "N genes"),
#       lty=c(1,0), pch=c(NA, 16), col=c("red3", "black"))

par(fig = c(0.75,1,0, 0.77), new = TRUE)
boxplot(samedates$HI, samedates$LO, col = c("gray", "red"), 
        #ylim = c(50,85), 
        axes = F, cex = 0.5)
mtext(side = 1,"High         Low", cex = 0.7)

#par(fig = c(0.77, 1, 0.6, 1), new = TRUE)
#legend("center", legend = c("High zone, upper limit of sand dollar bed", "Low zone, lower limit of sand dollar bed"),
#       col = c("black","red"))


```

# Computing whether tide is rising or falling
```{r}
# the lag function offsets the observation by one row!
hist(samedates$TideHeight - lag(samedates$TideHeight))

samedates <- samedates %>% 
  mutate(tidestatus = case_when(
    (TideHeight - lag(TideHeight)) > 0.015 ~ "rising", 
    (TideHeight - lag(TideHeight)) < -0.015 ~ "falling",
    (TideHeight - lag(TideHeight)) <= 0.015 & TideHeight - lag(TideHeight) >= -0.015 ~ "slack"
  ))


ggplot(samedates) + 
  aes(x = tidestatus, y = LO) + 
  geom_boxplot()


```

